---
title: "intro to sf"
---


```{r include = FALSE}
source("_setup.R")
```

---

class: inverse, center, middle

# `r nf()` Exercise

![:custom_hr]()

## `r ck()` ~40min

---

# Exercice

## Steps

1. Use `rgovcan` to download the atlas data ID = "f612e2b4-5c67-46dc-9a84-1154c649ab4e"
2. Load ‘data/dat.csv’ to access the data associated with the atlas
3. Create two objects containing data for Atlantic Puffin and Razorbill
4. Join the data for the Atlantic Puffin and Razorbill with the atlas grid
5. Create two convex hulls:
  - Atlantic Puffin with densities > 6 birds / km^2
  - Razorbill with densities > 1 birds / km^2
6. Do a zonal intersection between the convex hulls
7. Measure the overlapping area between the two convex hulls




---

# Exercice

## Solution

```{r exsf, include = TRUE, eval = FALSE, purl = TRUE}
# Download and import atlas grid
govcan_dl_resources("f612e2b4-5c67-46dc-9a84-1154c649ab4e", path = 'output')
unzip("output/AtlasGrid-GrilleAtlas.gdb.zip", exdir = 'output')
atlas <- st_read('output/AtlasGrid-GrilleAtlas.gdb', layer = 'AtlasGrid_GrilleAtlas')

# Import atlas data as csv
dat <- read.csv('./data/DensityData-DonneesDeDensite.csv')

# Join data with grid
atpu <- dat[dat$Group == 'ATPU', ] %>% group_by(Stratum) %>% summarize(ATPU = mean(Density))
razo <- dat[dat$Group == 'RAZO', ] %>% group_by(Stratum) %>% summarize(RAZO = mean(Density))
atlas <- left_join(atlas, atpu, by = c("id" = "Stratum")) %>%
         left_join(razo, by = c("id" = "Stratum"))


# Convex hulls
atpu_ch <- st_convex_hull(st_union(atlas[atlas$ATPU > 6, ]))
razo_ch <- st_convex_hull(st_union(atlas[atlas$RAZO > 1, ]))

# Intersection
int <- st_intersection(atpu_ch,razo_ch)

# Area
st_area(int)

# Visualize
plot(st_geometry(atlas))
plot(atpu_ch, border = '#b1646d', lwd = 2, add = TRUE)
plot(razo_ch, border = '#b1646d', lwd = 2, add = TRUE)
plot(int, col = '#074a6c', add = TRUE)

```
