---
title: "Exercise leaflet"
output:
  xaringan::moon_reader:
    css: [default, rd.css, rd-font.css, "hygge"]
    lib_dir: assets
    seal: false
    nature:
      highlightStyle: dracula
      countIncrementalSlides: false
      beforeInit: "macros.js"
---


```{r include = FALSE}
source("_setup.R")
```

---
class: inverse, center, middle

# `r nf()` Exercise

![:custom_hr]()

## `r ck()` ~40min


---
# Package [`leaflet`](https://cran.r-project.org/package=leaflet)


`r cdw(minutes = 40, seconds = 0)`


### .font90[Generate a leaflet map with:]

.font80[
- Two tile providers (hint: visit [this](http://leaflet-extras.github.io/leaflet-providers/preview/index.html) website)

- `hyp` and `acid`

- Set view to Service Canada Building (45 Alderney Dr. Dartmouth, NS B2Y 2N6)
  - Get coordinates from [Google Maps](https://www.google.ca/maps) or [OpenStreetMap](https://www.openstreetmap.org/)
]

### .font90[Use a custom color palette to render the rasters (hint: look up [`colorNumeric()`](https://www.rdocumentation.org/packages/leaflet/versions/2.0.3/topics/colorNumeric))]

### .font90[Render a single raster on map startup (hint: look up [`hideGroup()`](https://rdrr.io/cran/leaflet/man/showGroup.html))]


### .font90[Recreate the map you generated before with `leaflet` using `mapview`]

--

### .font90[Export it as a `html` file in the `output/` folder]

--

### .font90[Send it to your best friend!]



---

# Solution `leaflet`

.font90[
```{r leafletEx, include = FALSE, purl = FALSE}
# Colors
cols <- c("#c7cbce", "#687677", "#222d3d", "#25364a", "#172434",
          "#ad6a11", "#e6a331", "#e4be29", "#f2ea8b")

# Leaflet map
leaflet() %>%
  setView(lng = -63.569130, lat = 44.6652645, zoom = 5) %>%
  addProviderTiles('Esri.OceanBasemap', group = 'Ocean') %>%
  addProviderTiles("OpenTopoMap", group = "Topo") %>%
  addStarsImage(hyp, group = 'Hyp', project = FALSE,
                colors = colorNumeric(cols, hyp[[1]], na.color = "transparent")) %>%
  addStarsImage(acid, group = 'Acid', project = FALSE,
                colors = colorNumeric(cols, acid[[1]], na.color = "transparent")) %>%

  # Reset layer selection
  addLayersControl(baseGroups = c('Ocean', 'Topo'),
    overlayGroups = c('Hyp','Acid'),
    position = 'topleft') %>%

  # Hide 2 layers
  hideGroup(c("Fish","Acid"))
```
]


---

# Solution `mapview`

```{r mapviewEx, eval = FALSE, include = FALSE, purl = FALSE}
# Color ramp palette
pal <- colorRampPalette(cols)

# Mapview
mv <- mapview(hyp, col.regions = pal) +
      mapview(acid, col.regions = pal)

# Export
dir.create('output')
mapshot(mv, url = 'output/mapviewEx.html')

# Send email
# Not actually going to do this! But! If you want to send emails through R, check this blog post:
# https://blog.mailtrap.io/r-send-email/
```


